---
title: "UC data wrangling"
author: "Anupriya Kushwanshi"
date: "May 25, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

reading data
```{r}
library(tidyverse)

house <- read.table("C:/Users/anupr/Documents/Flex2/DAM/Hours-to-Pay-Mortgage.csv",sep = ",", header = T, col.names = c("City","State","Median Price","Mortgage Rate 30yr","Monthly Mortgage Payment","Median Income","Hours/month needed","Periods","Present Value", "X", "bin"))

tail(house)
house <- select(house, everything(), -c("X","bin"))
head(house)
    
```

Checking for null values
```{r}
colSums(is.na(house))
```
```{r}
str(house)
```
Changing the class of numerical variables
```{r}

  house$Median.Price <- as.integer(house$Median.Price)
  house$Median.Income <- as.factor(house$Median.Income)
  house$Monthly.Mortgage.Payment <- as.numeric(house$Monthly.Mortgage.Payment)
  house$Hours.month.needed <- as.numeric(house$Hours.month.needed)
  house$Mortgage.Rate.30yr <- house$Mortgage.Rate.30yr
```

Plots using subsetted data
```{r}

top_10_med_price <- head(arrange(house, by = Median.Price), 10)

str(top_10_med_price)
```

```{r}

top_10_med_price%>%
  ggplot(aes(x=reorder(City,-Median.Price), y=Median.Price, fill =  Median.Income ))+geom_bar(stat="identity")+xlab("City")+ylab("Median House Price")

```

```{r}
top_10_med_price%>%
  ggplot(aes(x=reorder(City,-Present.Value), y=Median.Price, fill = Monthly.Mortgage.Payment  ))+geom_bar(stat="identity")+xlab("City")+ylab("Median House Price")

```

Reading data in a faster way - good for big data
```{r}
library(data.table)

house.2 <- read.csv("C:/Users/anupr/Documents/Flex2/DAM/Hours-to-Pay-Mortgage.csv")
```
Read_csv and fread maintain white spaces. fread is fastest of all.
```{r}
house.3 <- read_csv("C:/Users/anupr/Documents/Flex2/DAM/Hours-to-Pay-Mortgage.csv")
system.time(house.1 <- fread("C:/Users/anupr/Documents/Flex2/DAM/Hours-to-Pay-Mortgage.csv"))
```

exporting data in multiple sheets in excel
```{r}
# install.packages("devtools")
'devtools::install_github("kassambara/r2excel")
library(r2excel)

multiple_df <- createWorkbook()
car_df <- createSheet(wb = multiple_df, sheetName = "Cars")
iris_df <- createSheet(wb = multiple_df, sheetName = "Iris")'
```

read from database
```{r}
'install.packages("RODBC")
require(RODBC)'
```
Qplot - plotting 2 types at once
```{r}
data("mpg")
head(mpg)
qplot(temperature, pressure, data = pressure, geom = c("line", "point"))

```
Barplot
```{r}
head(mpg)
str(mpg)
qplot(data = mpg, manufacturer)
ggplot(mpg, aes(manufacturer, fill=class))+geom_bar()
```

Histogram
```{r}
mpg$year <- as.factor(mpg$year)
str(mpg)
unique(mpg$year)
ggplot(mpg, aes(x=year, fill=class))+geom_bar()
```

Stem and leaf plot
```{r}
stem(mpg$displ)
```
Line Chart - changes over time
```{r}
mpg%>%
ggplot(aes(x=year, y=hwy))+ geom_line() +geom_point(aes(colour = model))+  geom_line(aes(group = model), colour = "grey50")
```

 case study - data cleaning
```{r}
data("who")
head(who)
```
```{r}
dim(who)
```

```{r}
str(who)
```
```{r}
unique(who$newrel_f2534)
```
```{r}
colnames(who)
```
 
 There are so many columns with similar data. We see there are some numbers in each such column. This can be count.
```{r}
who_gather <- gather(who,new_sp_m014:newrel_f65 ,key="key",value = "cases" , na.rm = TRUE)
head(arrange(who_gather, by=desc(cases)))

```
```{r}
count(who_gather,key)
```
```{r}
 who_sep <- separate(who_gather, key, c("new or old", "type", "sex_age"), sep='_')
head(who_sep)
```
split sex and age after the first character
```{r}
who_sep_1 <- separate(who_sep, sex_age, c("sex","age"), sep=1)
head(who_sep_1)
```
```{r}
final_who <- select(who_sep_1, everything(), -c("iso2","iso3"))
head(final_who)
```
```{r}
final_who <- na.omit(final_who)
str(final_who)

```

Visualising new who dataset
```{r}
final_who%>%
  ggplot(aes(year, fill=sex))+geom_bar()+theme(legend.position = "top")
```

```{r}
final_who%>%
  ggplot(aes(year, fill=age))+geom_bar()
```

 Top 10 cases are from India
```{r}
top_10_country <- head(arrange(final_who, desc(cases)), 10)
top_10_country%>%
  ggplot(aes(x=top_10_country$country, fill=top_10_country$type))+geom_bar()
```
Top 50 cases are from India and China
```{r}
top_50_country <- head(arrange(final_who, desc(cases)), 50)
top_50_country%>%
  ggplot(aes(x=top_50_country$country, fill=top_50_country$type))+geom_bar()
```
```{r}
top_50_country <- head(arrange(final_who, desc(cases)), 50)
top_50_country%>%
  ggplot(aes(x=top_50_country$country, fill=top_50_country$sex))+geom_bar()
```

Bottom 50 
```{r}
bottom_50_country <- head(arrange(final_who), 50)
bottom_50_country%>%
  ggplot(aes(x=country, fill=sex))+geom_bar()
```

